\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage{fancyhdr}
\usepackage{xcolor}
\usepackage{tcolorbox}
\usepackage{hyperref}
\usepackage{caption}
\usepackage{circuitikz}

% Boolean to show/hide results
\newif\ifshowresults
\showresultstrue  % Set to \showresultsfalse to hide solutions

% Environment for solutions
\usepackage{comment}
\ifshowresults
    \newenvironment{solution}
    {\color{blue}\textbf{Solution:}\\}
    {}
\else
    \excludecomment{solution}
\fi

\pagestyle{fancy}
\fancyhf{}
\lhead{ECEN 222 - Electronic Circuits}
\rhead{Spring 2026}
\cfoot{\thepage}

\begin{document}

\begin{center}
    {\LARGE \textbf{Assignment 05}}\\
    {\LARGE \textbf{Operational Amplifier Applications}}\\
    \vspace{0.3cm}
    {\large ECEN 222, Spring 2026}\\
    \vspace{0.2cm}
    {\large University of Nebraska-Lincoln}\\
    \vspace{0.5cm}
\end{center}

\section*{Overview}

This assignment focuses on operational amplifier circuits and their applications. You will analyze various op-amp configurations using pencil-and-paper calculations, then verify your results using Multisim simulation. Understanding op-amp circuits is fundamental to analog circuit design, signal processing, and instrumentation.

\vspace{0.3cm}

\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=\textbf{Unipolar Power Supply Configuration}]
\textbf{All circuits in this assignment use UNIPOLAR power supplies} (not dual supplies):
\begin{itemize}
    \item Power supply voltage: $V_{DD} = 15$ V (positive supply)
    \item Ground (GND): 0 V
    \item \textbf{AC Ground (Reference Voltage):} $V_{ref} = \frac{V_{DD}}{2} = 7.5$ V
\end{itemize}



In Multisim, you must connect a 7.5 V DC voltage source to these nodes. This provides the DC bias point that allows the op-amp to swing both 'positive' and 'negative' around 7.5 V with a unipolar supply.
\end{tcolorbox}

\vspace{0.3cm}

\section*{Submission Requirements}

Submit a \textbf{single PDF document} containing:

\begin{itemize}
    \item All hand-written or typed calculations with clear steps shown
    \item Written explanations and assumptions
    \item Multisim screenshots with component values and measurements visible
    \item Comparison of calculated vs. simulated results with percent error
\end{itemize}

\vspace{0.3cm}

\section*{Software Requirements}

\begin{itemize}
    \item \textbf{Multisim:} Available through the NI Academic Site License. Download from \url{https://www.ni.com/} or use computers in the lab.
\end{itemize}

\newpage

\section*{Part 1: Inverting Amplifier Analysis [30 points]}

Analyze the following inverting amplifier circuit.

\subsection*{Circuit Specifications:}

\begin{itemize}
    \item Op-amp: LM741 (or similar general-purpose op-amp)
    \item Input resistor: $R_1 = 10$ k$\Omega$
    \item Feedback resistor: $R_f = 47$ k$\Omega$
    \item Input voltage: $v_{in} = 8.0$ V DC (referenced to AC ground as +0.5V, since 8 - 7.5 V = 0.5 V)
    \item Power supply: $V_{DD} = 15$ V, GND = 0 V
    \item AC ground (reference): $V_{ref} = 7.5$ V (at non-inverting input)
\end{itemize}

\begin{center}
\begin{circuitikz}[scale=1.0, every node/.style={font=\small}]
    % Op-amp
    \node[op amp, scale=1.0] (opamp) at (3,0) {};

    % Input
    \coordinate (vin) at ($(opamp.-)+(-3.0,0)$);
    \draw (vin) node[left] {$v_{in}$}
        to[R, l=$R_1$, o-] (opamp.-);

    % Non-inverting input to AC ground (Vref = 7.5V)
    \coordinate (vref) at ($(opamp.+)+(0,-1.0)$);
    \draw (opamp.+) -- (vref);
    \draw (vref) node[ground] {};
    \node[right, red] at ($(vref)+(0.3,-0.2)$) {$V_{ref}=7.5$V};

    % Output
    \draw (opamp.out) to[short, -o] (5.2,0) node[right] {$v_{out}$};

    % Feedback resistor
    \draw (opamp.out) -- ++(0,1.6)
        to[R, l=$R_f$] ++(-3.0,0)
        |- (opamp.-);

\end{circuitikz}
\captionof{figure}{Inverting amplifier circuit (AC ground at $V_{ref}=7.5$ V)}
\end{center}

\subsection*{Analysis Tasks:}

\begin{enumerate}[label=\textbf{1.\arabic*}]
    \item \textbf{Theoretical Analysis [15 points]:} Using pencil and paper, calculate:
    \begin{itemize}
        \item The voltage gain $A_v = \frac{v_{out}}{v_{in}}$
        \item The input impedance $R_{in}$
        \item The output voltage $v_{out}$ for the given input
    \end{itemize}
    
\begin{solution}
\textbf{Voltage Gain:}

For an inverting amplifier, the gain is:
\begin{align*}
A_v &= -\frac{R_f}{R_1} = -\frac{47\text{ k}\Omega}{10\text{ k}\Omega} = -4.7
\end{align*}

\textbf{Input Impedance:}

The input impedance seen by the source is simply $R_1$ (since the inverting input is at virtual ground):
\begin{align*}
R_{in} = R_1 = 10\text{ k}\Omega
\end{align*}

\textbf{Output Voltage:}

The input voltage referenced to AC ground is:
\begin{align*}
v_{in,ac} &= v_{in} - V_{ref} = 8.0 - 7.5 = 0.5\text{ V}
\end{align*}

The output referenced to AC ground:
\begin{align*}
v_{out,ac} &= A_v \cdot v_{in,ac} = -4.7 \times 0.5 = -2.35\text{ V}
\end{align*}

The absolute output voltage (referenced to GND):
\begin{align*}
v_{out} &= V_{ref} + v_{out,ac} = 7.5 + (-2.35) = 5.15\text{ V}
\end{align*}
\end{solution}

\vspace{1cm}
    
    \item \textbf{Assumptions and Explanation [10 points]:} In a written paragraph, explain:
    \begin{itemize}
        \item What assumptions you made about the op-amp (list all ideal op-amp assumptions)
        \item Why the circuit is called an "inverting" amplifier
        \item How the virtual ground concept applies to this circuit (note: the inverting input will be at 7.5 V, not 0 V)
        \item Why we need the AC ground reference at 7.5 V for unipolar supply operation
        \item What would happen if $R_f < R_1$ (would it still be an amplifier?)
    \end{itemize}
    
\begin{solution}
\textbf{Ideal Op-Amp Assumptions:}
\begin{enumerate}
    \item Infinite input impedance (no current flows into op-amp inputs)
    \item Zero output impedance (can source/sink any current)
    \item Infinite open-loop gain ($A_{OL} \to \infty$)
    \item Infinite bandwidth (gain constant at all frequencies)
    \item Zero offset voltage (when inputs equal, output is zero)
\end{enumerate}

\textbf{Why "Inverting":} The circuit is called inverting because the output signal is 180° out of phase with the input signal. When the input increases above $V_{ref}$, the output decreases below $V_{ref}$, and vice versa. The negative sign in $A_v = -4.7$ indicates this phase inversion.

\textbf{Virtual Ground Concept:} In this circuit, the non-inverting input is connected to $V_{ref} = 7.5$ V. Due to negative feedback and infinite gain, the inverting input is forced to the same voltage as the non-inverting input, creating a "virtual AC ground" at 7.5 V. This is NOT 0 V (true ground), but it acts as a reference point that remains constant as signals vary around it.

\textbf{Why AC Ground at 7.5 V:} With a unipolar supply (0 V to 15 V), the op-amp can only output voltages within this range. Setting the reference at $V_{ref} = V_{DD}/2 = 7.5$ V allows the output to swing both "positive" (toward 15 V) and "negative" (toward 0 V) around this midpoint, maximizing the dynamic range.

\textbf{If $R_f < R_1$:} If $R_f < R_1$, the magnitude of the gain would be less than 1 ($|A_v| < 1$). The circuit would still invert the signal, but it would attenuate rather than amplify. For example, if $R_f = 5$ k$\Omega$ and $R_1 = 10$ k$\Omega$, then $A_v = -0.5$ — this is an inverting attenuator, not an amplifier.
\end{solution}

\vspace{1cm}
    
    \item \textbf{Design Question [5 points]:} If you needed a gain of $A_v = -20$ and wanted to keep $R_1 = 10$ k$\Omega$, what value should $R_f$ be? Show your calculation.

\begin{solution}
Using the gain formula for an inverting amplifier:
\begin{align*}
A_v &= -\frac{R_f}{R_1}\\
-20 &= -\frac{R_f}{10\text{ k}\Omega}\\
R_f &= 20 \times 10\text{ k}\Omega = 200\text{ k}\Omega
\end{align*}

\textbf{Answer:} $R_f = 200$ k$\Omega$
\end{solution}
\end{enumerate}

\newpage

\section*{Part 2: Integrator Circuit Analysis [30 points]}

Analyze the following inverting integrator circuit.

\subsection*{Circuit Specifications:}

\begin{itemize}
    \item Input resistor: $R = 100$ k$\Omega$
    \item Feedback capacitor: $C = 1$ $\mu$F
    \item Input voltage: $v_{in} = 9.0$ V DC (constant, referenced to AC ground)
    \item Power supply: $V_{DD} = 15$ V, GND = 0 V
    \item AC ground (reference): $V_{ref} = 7.5$ V (at non-inverting input)
    \item Initial condition: Assume $v_{out}(0) = 7.5$ V (at AC ground)
\end{itemize}

\begin{center}
\begin{circuitikz}[scale=1.0, every node/.style={font=\small}]
    % Op-amp
    \node[op amp, scale=1.0] (opamp) at (3.2,0) {};

    % Input resistor
    \coordinate (invL) at (2.0,0 |- opamp.-);
    \draw (0.3,0 |- opamp.-) node[left] {$v_{in}$}
        to[R, l=$R$, o-] (invL) -- (opamp.-);

    % Non-inverting input to AC ground
    \coordinate (vref) at ($(opamp.+)+(0,-1.2)$);
    \draw (opamp.+) -- (vref);
    \draw (vref) node[ground] {};
    \node[right, red] at ($(vref)+(0.3,-0.2)$) {$V_{ref}=7.5$V};

    % Feedback capacitor
    \coordinate (fbTop) at ($(opamp.out)+(0,1.4)$);
    \draw (opamp.out) -- (fbTop)
        to[C, l^=$C$] ++(-2.0,0)
        -| (opamp.-);

    % Output
    \draw (opamp.out) to[short, -o] (4.7,0) node[right] {$v_{out}$};

\end{circuitikz}
\captionof{figure}{Inverting integrator circuit (AC ground at $V_{ref}=7.5$ V)}
\end{center}

\subsection*{Analysis Tasks:}

\begin{enumerate}[label=\textbf{2.\arabic*}]
    \item \textbf{Theoretical Analysis [15 points]:} Using pencil and paper, calculate:
    \begin{itemize}
        \item The time constant $\tau = RC$
        \item The output voltage as a function of time: $v_{out}(t) = -\frac{1}{RC} \int_0^t (v_{in} - V_{ref}) d\tau + v_{out}(0)$
        \item Note: Since $v_{in}$ is constant at 9.0 V and $V_{ref} = 7.5$ V, the effective input is $(v_{in} - V_{ref}) = 1.5$ V
        \item Calculate $v_{out}$ at $t = 0.05$ s and $t = 0.1$ s
        \item How long until the output reaches 5.0 V (referenced to GND, not to $V_{ref}$)?
    \end{itemize}
    
\begin{solution}
\textbf{Time Constant:}
\begin{align*}
\tau = RC = (100 \times 10^3\,\Omega)(1 \times 10^{-6}\text{ F}) = 0.1\text{ s}
\end{align*}

\textbf{Output Voltage Function:}

The effective input relative to AC ground is:
\begin{align*}
v_{in, eff} = v_{in} - V_{ref} = 9.0 - 7.5 = 1.5\text{ V}
\end{align*}

For a constant input, the integrator output is:
\begin{align*}
v_{out}(t) &= -\frac{1}{RC} \int_0^t v_{in, eff} \, d\tau + v_{out}(0)\\
&= -\frac{1}{RC} \cdot v_{in, eff} \cdot t + v_{out}(0)\\
&= -\frac{1.5}{0.1} \cdot t + 7.5\\
&= -15t + 7.5\text{ V}
\end{align*}

\textbf{At $t = 0.05$ s:}
\begin{align*}
v_{out}(0.05) = -15(0.05) + 7.5 = -0.75 + 7.5 = 6.75\text{ V}
\end{align*}

\textbf{At $t = 0.1$ s:}
\begin{align*}
v_{out}(0.1) = -15(0.1) + 7.5 = -1.5 + 7.5 = 6.0\text{ V}
\end{align*}

\textbf{Time to reach 5.0 V:}
\begin{align*}
5.0 &= -15t + 7.5\\
15t &= 2.5\\
t &= \frac{2.5}{15} = 0.167\text{ s} = 167\text{ ms}
\end{align*}
\end{solution}

\vspace{1cm}
    
    \item \textbf{Circuit Operation Explanation [15 points]:} Explain:
    \begin{itemize}
        \item How the integrator works (describe the relationship between capacitor current and output voltage)
        \item Why the output voltage changes over time even with a constant DC input
        \item What the negative sign in the transfer function means
        \item Why we use $(v_{in} - V_{ref})$ instead of just $v_{in}$ in our calculations
        \item What would happen if you applied a step input (instead of DC) to an integrator
    \end{itemize}
    
\begin{solution}
\textbf{How the Integrator Works:}

The inverting input is at virtual AC ground (7.5 V). Current flows through $R$ from the input: $i = \frac{v_{in} - V_{ref}}{R}$. This current cannot flow into the op-amp (infinite input impedance), so it must flow into the capacitor. The capacitor equation is $i = C \frac{dv}{dt}$, which means $v_{out} = -\frac{1}{C}\int i \, dt$. Combining these gives the integrator transfer function.

\textbf{Why Output Changes with Constant DC Input:}

Even though the input is constant DC, there is a constant current flowing through $R$ into the capacitor. This constant current continuously charges (or discharges) the capacitor, causing the voltage across it to change linearly over time. The integrator integrates the DC offset relative to $V_{ref}$, resulting in a ramp output. In an ideal integrator with DC input, the output would eventually saturate at the power supply rail.

\textbf{Negative Sign Meaning:}

The negative sign indicates phase inversion. When the input is above $V_{ref}$ (positive), current flows into the capacitor's negative plate, causing the output voltage to decrease (ramp downward). Conversely, when input is below $V_{ref}$, the output ramps upward. This is characteristic of inverting op-amp configurations.

\textbf{Why $(v_{in} - V_{ref})$:}

We use $(v_{in} - V_{ref})$ because the inverting input is held at $V_{ref}$ by the virtual ground principle. Only the voltage difference drives current through $R$. If $v_{in} = V_{ref} = 7.5$ V, then $(v_{in} - V_{ref}) = 0$, no current flows, and the output remains constant. The circuit integrates relative to the AC ground reference.

\textbf{Step Input Response:}

If a step input is applied, the integrator produces a ramp output. For a positive step (sudden increase), the output ramps down linearly. For a negative step (sudden decrease), the output ramps up linearly. The slope of the ramp is determined by $-\frac{\Delta V_{in}}{RC}$. This is why integrators are used in waveform generation (e.g., triangle waves from square waves).
\end{solution}

\vspace{1cm}
    
    \item \textbf{Frequency Response [BONUS: 5 points]:} The integrator's frequency domain transfer function is $H(j\omega) = -\frac{1}{j\omega RC}$. Calculate the magnitude of the gain $|H(f)|$ at $f = 10$ Hz. Show your work and explain in 1-2 sentences what this tells you about the integrator's behavior at different frequencies.

\begin{solution}
The magnitude of the transfer function is:
\begin{align*}
|H(j\omega)| &= \left|-\frac{1}{j\omega RC}\right| = \frac{1}{\omega RC} = \frac{1}{2\pi f RC}
\end{align*}

At $f = 10$ Hz with $R = 100$ k$\Omega$ and $C = 1$ $\mu$F:
\begin{align*}
|H(10\text{ Hz})| &= \frac{1}{2\pi (10)(100 \times 10^3)(1 \times 10^{-6})}\\
&= \frac{1}{2\pi (10)(0.1)}\\
&= \frac{1}{2\pi} = \frac{1}{6.283} \approx 0.159\text{ or } 15.9\%
\end{align*}

\textbf{Interpretation:} The integrator has a gain that decreases with increasing frequency ($-20$ dB/decade slope). At low frequencies, the gain is high, amplifying low-frequency signals and DC offsets significantly. At high frequencies, the gain is low, attenuating high-frequency components. This makes the integrator act as a low-pass filter.
\end{solution}
\end{enumerate}

\newpage

\section*{Part 3: Summing Amplifier Analysis [40 points]}

Analyze the following three-input summing amplifier circuit.

\subsection*{Circuit Specifications:}

\begin{itemize}
    \item Input resistors: $R_1 = 10$ k$\Omega$, $R_2 = 20$ k$\Omega$, $R_3 = 5$ k$\Omega$
    \item Feedback resistor: $R_f = 20$ k$\Omega$
    \item Input voltages: $v_1 = 8.5$ V DC, $v_2 = 8.0$ V DC, $v_3 = 7.0$ V DC
    \item Power supply: $V_{DD} = 15$ V, GND = 0 V
    \item AC ground (reference): $V_{ref} = 7.5$ V (at non-inverting input)
\end{itemize}

\begin{center}
\begin{circuitikz}[scale=0.9, every node/.style={font=\small}]
    % Op-amp (centered)
    \node[op amp, scale=1.0] (opamp) at (3.2,0) {};

    \coordinate (sum) at (1.5,0 |- opamp.-);

    % Inputs and resistors
    \draw ([yshift=1.2cm]-1.5, 0 |- opamp.-) node[left] {$v_1$} to[R, l=$R_1$, o-] ([yshift=1.2cm]sum);
    \draw (-1.5, 0 |- opamp.-) node[left] {$v_2$} to[R, l=$R_2$, o-] (sum);
    \draw ([yshift=-1.2cm]-1.5, 0 |- opamp.-) node[left] {$v_3$} to[R, l=$R_3$, o-] ([yshift=-1.2cm]sum);

    % Vertical bus + junction dot
    \draw ([yshift=1.2cm]sum) -- ([yshift=-1.2cm]sum);
    \fill (sum) circle (1.2pt);

    % Connect to inverting input
    \draw (sum) -- (opamp.-);

    % Non-inverting input to AC ground
    \coordinate (vref) at ($(opamp.+)+(0,-1.2)$);
    \draw (opamp.+) -- (vref);
    \draw (vref) node[ground] {};
    \node[right, red] at ($(vref)+(0.3,-0.2)$) {$V_{ref}=7.5$V};

    % Output
    \draw (opamp.out) to[short, -o] (5.7,0) node[right] {$v_{out}$};

    % Feedback resistor
    \draw (opamp.out) -- ++(0,1.4)
        to[R, l^=$R_f$] ++(-2.0,0)
        -| (opamp.-);

\end{circuitikz}
\captionof{figure}{Three-input summing amplifier (AC ground at $V_{ref}=7.5$ V)}
\end{center}

\subsection*{Analysis Tasks:}

\begin{enumerate}[label=\textbf{3.\arabic*}]
    \item \textbf{Theoretical Analysis [20 points]:} Using pencil and paper, calculate:
    \begin{itemize}
        \item The weight (gain coefficient) for each input: $w_1 = \frac{R_f}{R_1}$, $w_2 = \frac{R_f}{R_2}$, $w_3 = \frac{R_f}{R_3}$
        \item The output voltage using: $v_{out} = -\left(\frac{R_f}{R_1}v_1 + \frac{R_f}{R_2}v_2 + \frac{R_f}{R_3}v_3\right)$
        \item Verify your calculation by substituting the given input voltages
    \end{itemize}
    
\begin{solution}
\textbf{Weight Coefficients:}
\begin{align*}
w_1 &= \frac{R_f}{R_1} = \frac{20\text{ k}\Omega}{10\text{ k}\Omega} = 2.0\\
w_2 &= \frac{R_f}{R_2} = \frac{20\text{ k}\Omega}{20\text{ k}\Omega} = 1.0\\
w_3 &= \frac{R_f}{R_3} = \frac{20\text{ k}\Omega}{5\text{ k}\Omega} = 4.0
\end{align*}

\textbf{Output Voltage Calculation:}

First, convert inputs to AC voltages (relative to $V_{ref} = 7.5$ V):
\begin{align*}
v_{1,ac} &= v_1 - V_{ref} = 8.5 - 7.5 = 1.0\text{ V}\\
v_{2,ac} &= v_2 - V_{ref} = 8.0 - 7.5 = 0.5\text{ V}\\
v_{3,ac} &= v_3 - V_{ref} = 7.0 - 7.5 = -0.5\text{ V}
\end{align*}

The output relative to $V_{ref}$:
\begin{align*}
v_{out,ac} &= -(w_1 v_{1,ac} + w_2 v_{2,ac} + w_3 v_{3,ac})\\
&= -(2.0 \times 1.0 + 1.0 \times 0.5 + 4.0 \times (-0.5))\\
&= -(2.0 + 0.5 - 2.0)\\
&= -0.5\text{ V}
\end{align*}

The absolute output voltage (referenced to GND):
\begin{align*}
v_{out} &= V_{ref} + v_{out,ac} = 7.5 + (-0.5) = 7.0\text{ V}
\end{align*}

\textbf{Verification using absolute voltages:}
\begin{align*}
v_{out} &= V_{ref} - \left(\frac{R_f}{R_1}(v_1 - V_{ref}) + \frac{R_f}{R_2}(v_2 - V_{ref}) + \frac{R_f}{R_3}(v_3 - V_{ref})\right)\\
&= 7.5 - \left(2.0(1.0) + 1.0(0.5) + 4.0(-0.5)\right)\\
&= 7.5 - 0.5 = 7.0\text{ V} \quad \checkmark
\end{align*}
\end{solution}

\vspace{1cm}
    
    \item \textbf{Circuit Operation Explanation [15 points]:} Explain:
    \begin{itemize}
        \item How the summing amplifier works (describe the current flow from each input)
        \item Why the inverting input is at virtual AC ground (7.5 V)
        \item Why the output is inverted and weighted
        \item What happens if all three inputs are at $V_{ref}$ (7.5 V) - what would the output be?
        \item Give one practical application where a summing amplifier would be useful
    \end{itemize}
    
\begin{solution}
\textbf{How the Summing Amplifier Works:}

Current flows from each input through its respective resistor toward the inverting input (virtual ground at 7.5 V). The current from input $i$ is $i_i = \frac{v_i - V_{ref}}{R_i}$. Since no current flows into the op-amp (infinite input impedance), all input currents sum together and flow through the feedback resistor $R_f$. By KCL at the inverting node: $i_1 + i_2 + i_3 = i_f$. The voltage across $R_f$ creates the output: $v_{out} - V_{ref} = -i_f R_f$.

\textbf{Virtual AC Ground:}

The non-inverting input is connected to $V_{ref} = 7.5$ V. Due to negative feedback and the op-amp's infinite gain, the inverting input is forced to match this voltage. This creates a "virtual AC ground" at 7.5 V. It's called virtual because no current flows into the op-amp terminal, yet the voltage remains fixed at 7.5 V, serving as a stable reference point for the circuit.

\textbf{Inverted and Weighted Output:}

The output is inverted because current flowing into the inverting node (from positive input voltages) must flow out through $R_f$, pulling the output below $V_{ref}$. Each input is weighted by its ratio $\frac{R_f}{R_i}$, meaning inputs with smaller resistors have larger weights (contribute more to the output). This allows precise control of how much each signal contributes to the sum.

\textbf{All Inputs at $V_{ref}$:}

If $v_1 = v_2 = v_3 = V_{ref} = 7.5$ V, then $(v_i - V_{ref}) = 0$ for all inputs. No current flows through any input resistor, so $i_f = 0$, and $v_{out} = V_{ref} = 7.5$ V. The output remains at the AC ground reference.

\textbf{Practical Application:}

Audio mixing consoles use summing amplifiers to combine multiple audio signals (microphones, instruments) into a single output channel. Each input can be weighted differently (volume control) by choosing appropriate resistor values, and the mixed signal can be inverted back if needed with another stage.
\end{solution}

\vspace{1cm}
    
    \item \textbf{Design Modification [5 points]:} If you wanted all three inputs to have equal weight (equal gain), what should the values of $R_1$, $R_2$, and $R_3$ be? State your answer and briefly explain your reasoning.

\begin{solution}
For equal weights, we need:
\begin{align*}
w_1 = w_2 = w_3 \quad \Rightarrow \quad \frac{R_f}{R_1} = \frac{R_f}{R_2} = \frac{R_f}{R_3}
\end{align*}

This requires:
\begin{align*}
R_1 = R_2 = R_3
\end{align*}

\textbf{Answer:} All three input resistors should be equal. A practical choice would be $R_1 = R_2 = R_3 = 20$ k$\Omega$ (equal to $R_f$), which gives each input a weight of $w = 1.0$.

\textbf{Reasoning:} Since the weight for each input is $w_i = \frac{R_f}{R_i}$, making all input resistors equal ensures all weights are equal. Choosing them equal to $R_f$ gives unity weight, meaning each input contributes equally to the inverted sum. For example, if $R_1 = R_2 = R_3 = R_f = 20$ k$\Omega$, then $v_{out} = -(v_1 + v_2 + v_3)$ (referenced to AC ground).
\end{solution}
\end{enumerate}

\newpage

\section*{Part 4: Multisim Simulation and Verification [Bonus: 20 points]}

Choose \textbf{ONE} of the three circuits from Parts 1-3 to build and simulate in Multisim.

\subsection*{Simulation Tasks:}

\begin{enumerate}[label=\textbf{4.\arabic*}]
    \item \textbf{Circuit Construction [5 points]:} Build your chosen circuit in Multisim showing:
    \begin{itemize}
        \item Op-amp (use LM741 or similar from the library)
        \item All resistors (and capacitor for integrator) with correct values clearly labeled
        \item DC voltage source(s) for input(s)
        \item \textbf{Unipolar power supply:} $V_{DD} = 15$ V to $V+$ pin, GND (0 V) to $V-$ pin
        \item \textbf{AC ground source:} 7.5 V DC source connected to the non-inverting (+) input
        \item Multimeter or probe at the output to measure $v_{out}$
    \end{itemize}
    Include a clear screenshot of your complete circuit.
    
    \item \textbf{DC Operating Point Analysis [10 points]:} 
    \begin{itemize}
        \item Run a DC Operating Point analysis or use multimeters to measure the output voltage
        \item Capture a screenshot showing the measured output voltage
        \item Compare your simulated result to your hand-calculated theoretical value
        \item Calculate the percent error.
    \end{itemize}
    
    \item \textbf{Analysis and Discussion [5 points]:} In a short paragraph, discuss:
    \begin{itemize}
        \item How close your simulation matched your calculations
        \item Possible reasons for any discrepancy (real vs. ideal op-amp)
        \item What you learned from building and simulating the circuit
        \item Any challenges you encountered in Multisim
    \end{itemize}
\end{enumerate}

\begin{solution}
\textbf{Expected Simulation Results:}

\begin{itemize}
    \item \textbf{Part 1 (Inverting Amplifier):} The simulated output should be very close to the calculated 5.15 V. Real op-amps have finite gain (typically 100,000 for LM741), which may cause a small error (< 1\%). The output may also show a slight offset voltage (typically ± few mV).
    
    \item \textbf{Part 2 (Integrator):} In transient analysis, the output should ramp down linearly from 7.5 V at a rate of -15 V/s. At $t = 0.05$ s, expect $\approx 6.75$ V; at $t = 0.1$ s, expect $\approx 6.0$ V. In practice, the integrator may saturate toward the lower rail if run too long. Real op-amps have input bias currents that can cause drift.
    
    \item \textbf{Part 3 (Summing Amplifier):} The simulated output should be very close to 7.0 V. Small discrepancies (< 1\%) can arise from resistor tolerances (use exact values in simulation) and op-amp imperfections (finite gain, offset voltage).
\end{itemize}

\textbf{Common Discrepancies:}
\begin{itemize}
    \item Finite open-loop gain causes small gain errors.
    \item Input offset voltage (1-5 mV for LM741) shifts the output slightly.
    \item Input bias current causes voltage drops across high-value resistors.
    \item Slew rate limitations affect transient response (not DC steady-state).
\end{itemize}

\textbf{Key Learning Points:}
Students should observe that real op-amps behave very close to ideal for these low-frequency DC circuits. The virtual ground concept can be verified by probing the inverting input (should measure 7.5 V). Understanding the unipolar supply configuration is crucial for single-supply designs.
\end{solution}

\end{document}
